\input{preamble.tex}

\input{commands.tex}

\title{PhD proposal document}
\author{Ioannis Paraskevakos}

\begin{document}
\maketitle
\pagebreak

\section{Introduction}
As scientific simulations achieve new scaling milestones in terms of simulation time and granularity,
they start to produce order of Gigabytes to Terabytes data that need to be analyzed. Based on that
analysis new simulations can be setup and executed. This data-analysis driven simulations can lead to
better more accurate results. Providing this style of execution requires first the ability to co-locate
MPI style simulations with data analytics.

The PST model proposed [cite here from Vivek] provides a unified way to express any type of workflow. 
Until now it is used mainly for abstracting scientific simulation pipelines and it successfully used
to drive this type of applications. On top of PST's implementation, Ensemble Toolkit, several abstractions
are defined that describe complex scientific workflows. Despite this fact, there are no abstractions
that allow to argue over data and define complex data analytics workflows. 

I propose a set of abstractions on the middleware layer, as well as, the application layer that allow
to define data analytics on HPC system and extend the PST model to be able to support data analysis
driven simulations.


\section{Related Work}
\section{Definitions}

\section{RMSD MPI4Py Implementation}
\section{Comparing the execution time with different frameworks}

\bibliographystyle{alpha}
\bibliography{sample}
\end{document}