The SelfLet framework~\cite{bindelli2008building} is an autonomic software system. 
This system provides a set of autonomic components, called SelfLets, that operate 
in order to achieve a goal. Each SelfLet provides a set of services, behaviors 
and policies. A SelfLet can be part of a network which allows it to utilize 
services from other SelfLets to achieve its goal. A SelfLet system has been used 
in a distributed sense to achieve load balanced service requests~\cite{calcavecchia2010emergence}.

The DIOS++ framework~\cite{liu2003dios} offers a rule-based autonomic management 
system for scientific applications. DIOS++ provides abstractions to create sensor 
and actuator which allow runtime monitoring and control, a distributed network to 
connect and manage the sensor/actuators and a distributed engine to execute parts 
of the application based on user defined rules.

Cloud4IoT~\cite{pizzolli2016cloud4iot} is an autonomic system which allows automatic 
deployment and configuration of data-intensive workloads on cloud and edge devices, 
and Internet-of-Things (IoT) application support. Cloud4IoT proposes an architecture 
where a Cloud is used as a central entity of computation, Edge devices are used to 
execute the initial processing steps of a data-intensive application, and IoT 
gateways are used as the gateways for sensors to connect to the system and 
transmit data.

CometCloud~\cite{diazmontes2015cometcloud} is an autonomic software system that 
enables scientific applications on a federation of resources. It has been used 
to enable simulation and data-intensive applications on heterogeneous resources, 
HPC and Cloud, that consumed million of core hours.

Pandey et al.~\cite{pandey2012autonomic} are proposing an autonomic cloud 
environment for executing multi user Electrocardiogram (ECG) data analysis 
workflows. The authors propose a three layer architecture. The ECG analysis 
software is the top layer of the architecture. The second layer contains the a 
scaling manager, a workflow engine, and a workload distribution engine. The cloud 
infrastructure and a authentication mechanism create the third layer of the 
architecture. All three layers are provided as a Service.